---
title: "Analysis for x linguistic processing"
output: html_notebook
---

```{r}
shhh <- suppressPackageStartupMessages # It's a library, so shhh!

shhh(library( mgcv ))
shhh(library(dplyr))
shhh(library(ggplot2))
shhh(library(lme4))
shhh(library(tidymv))
shhh(library(gamlss))
shhh(library(gsubfn))
shhh(library(lmerTest))
shhh(library(tidyverse))
shhh(library(boot))
shhh(library(rsample))
shhh(library(plotrix))
shhh(library(ggrepel))
shhh(library(mgcv))
library(jmuOutlier) # For paired permutation tests

options(JULIA_HOME = "/Applications/Julia-1.8.app/Contents/Resources/julia/bin/")
library(jglmm)
jglmm_setup()

theme_set(theme_bw())
options(digits=4)
options(dplyr.summarise.inform = FALSE)
```

```{r}

set.seed(444)
langs = c("du", "en", "fi", "ge", "gr", "he", "it", "ko", "sp", "tr", "ru")
psychometrics = c("total_rt", "gaze_rt", "firstfix_rt")
models = c("mgpt_sc", "mgpt_lc", "monot_30m", "monot_all")
comps = c("baseline", "add", "replace")

```

## Compute DLL for Each Language

```{r}

model_cross_val = function(form, df, d_var, mixed_effects, num_folds=10){
  
  folds <- cut(seq(1,nrow(df)),breaks=num_folds,labels=FALSE)
  
  estimates <- c()
  models <- c()
  for(i in 1:num_folds){
    testIndexes = which(folds==i,arr.ind=TRUE)
    testData = df[testIndexes,]
    trainData = df[-testIndexes,]

    if(mixed_effects){
      #model = lmer(as.formula(form), data = trainData)
      model = jglmm(as.formula(form), data = trainData)

    } else {
      model = lm(as.formula(form), data = trainData)
    }

    stdev = sigma(model)
    densities <- log(dnorm(testData[[d_var]],
                          mean=predict(model, newdata=testData),
                          sd=stdev))

    estimates <- c(estimates, densities)
  }

  return(estimates)
}

```


## Replace Surprisal w/ Entropy

Here "bl" = baseline with only surprisal and "0" = model where surprisal has been replaced with entropy at the current word

```{r output=FALSE, message=FALSE}

regression_names = c("baseline", "add", "replace")

dll_raw_df = data.frame()
for (lang in langs) {

  print(paste0("Fitting model for ", lang))
  
  df = read.csv(paste0("../data/merged_data/", lang, ".csv"))
  
  for (m in models) {
    
    df_eval = df %>% filter(model == m) %>%
      drop_na()
  
  for (psychometric in psychometrics) {
    
    regression_forms = c(
      paste0(psychometric, " ~ surp + prev_surp + prev2_surp + freq + len + prev_freq + prev_len + prev2_freq + prev2_len"),
      paste0(psychometric, " ~ ent + prev_ent + prev2_ent + surp + prev_surp + prev2_surp + freq + len + prev_freq + prev_len + prev2_freq + prev2_len"),
      paste0(psychometric, " ~ ent + prev_ent + prev2_ent + freq + len + prev_freq + prev_len + prev2_freq + prev2_len")
    )
    
    loglik_df = data.frame(names=regression_names, forms=regression_forms) %>%
      mutate(logliks = map(regression_forms, model_cross_val, df=df_eval, d_var=psychometric, mixed_effects = F )) %>%
      dplyr::select(-forms)
    
    loglik_df = loglik_df %>% unnest(cols = c(logliks)) %>% mutate(lang = lang, psychometric = psychometric, model = m)
    dll_raw_df = rbind(dll_raw_df, loglik_df)
    
  }
  }
}

```

## Individual language data

```{r}
comps = c("add", "replace")

dll_xlang_ent_df = data.frame()
for(l in langs){
  
  print(paste0("Tests for ", l)) 
  
  for (ps in psychometrics){
    for(c in comps){

      for(m in models) {
      if(c != "baseline") {
        target_df = dll_raw_df %>% filter(psychometric == ps, names == c, lang == l, model == m)
        baseline_df = dll_raw_df %>% filter(psychometric == ps, names == "baseline", lang == l, model == m)
        dll = target_df$logliks - baseline_df$logliks
        dll = dll[!is.na(dll)]
        ttest = perm.test(dll, num.sim = 500)
        dll_df = data.frame(comp = c, mean = mean(dll), upper = mean(dll) + (1.96 * std.error(dll)),
                            lower = mean(dll) - (1.96 * std.error(dll)), ttest_pval = ttest$p.value,
                            lang = l, psychometric = ps, model = m)
        dll_xlang_ent_df = rbind(dll_xlang_ent_df, dll_df)
      }
      }
    }
  }
}


```

## Data for languages as a whole

```{r}

dll_agg_ent_df = data.frame()
for (ps in psychometrics){
  for(c in comps){
    for(m in models){

      if(c != "baseline") {
        target_df = dll_raw_df %>% filter(psychometric == ps, names == c, model == m)
        baseline_df = dll_raw_df %>% filter(psychometric == ps, names == "baseline", model == m)
        dll = target_df$logliks - baseline_df$logliks
        dll = dll[!is.na(dll)]
        ttest = perm.test(dll, num.sim = 500)
        dll_df = data.frame(comp = c, mean = mean(dll), upper = mean(dll) + (1.96 * std.error(dll)),
                            lower = mean(dll) - (1.96 * std.error(dll)), ttest_pval = ttest$p.value,
                            lang = "All", psychometric = ps, model = m)
        dll_agg_ent_df = rbind(dll_agg_ent_df, dll_df)
      }
      }
  }
}


```



Plotting for these results

```{r}

# Merge the dataframes
dll_ent_plotting_df = rbind(dll_xlang_ent_df, dll_agg_ent_df)

dll_ent_plotting_df %>%
  mutate(lang = factor(lang, levels = c("All", "du", "en", "fi", "ge", "gr", "he", "it", "ko", "ru", "sp", "tr"),
       labels = c("All", "Dutch", "English", "Finnish", "German", "Greek", "Hebrew", "Italian", "Korean", "Russian",
                  "Spanish", "Turkish"))) %>%
  filter(model %in% c("mgpt_lc", "monot_all")) %>%
  rename(target = comp) %>%
  mutate(sig = case_when( ttest_pval >= 0.05 ~ " ",
                          ttest_pval < 0.05 & ttest_pval >= 0.01  ~ "*",
                          ttest_pval < 0.01 & ttest_pval >= 0.001  ~ "**",
                          ttest_pval < 0.001  ~ "***")) %>%
  mutate(psychometric = case_when(psychometric == "firstfix_rt" ~ "First Fixation", 
                                  psychometric == "gaze_rt" ~ "Gaze Duration", 
                                  psychometric == "total_rt" ~ "Total Fixation")) %>%
  mutate(model = if_else(model == "mgpt_lc", "mGPT", "monoT (all)")) %>%
  mutate(target = if_else(target == "add", "Add Entropy", "Replace Surprisal w/ Entropy")) %>%
  ggplot(aes(x = psychometric, y = mean, color = psychometric, shape = target, alpha=target)) +
    geom_hline(yintercept=0, color="black", linetype="dashed", alpha =0.5) +
    geom_point(position = position_dodge(width = 0.6) ) +
    geom_text(aes(y = if_else(target == "Add Entropy", 0.03, -0.067), label = sig), size = 3, show.legend = FALSE) +
    geom_errorbar(aes(ymin=lower, ymax=upper), width = 0.1, position = position_dodge(width = 0.6)) +
    ylab("Delta Log Likelihood (average per word)") + 
    xlab("") +
    facet_grid(model~lang) +
    scale_x_discrete(labels = c("FF", "GD", "TF")) +
    scale_alpha_discrete(range = c(1, 0.5)) +
    #scale_color_manual(values = c("#a1dab4", "#41b6c4", "#225ea8")) +
    #ggtitle("Contribution of Surprisal to ΔLL") + 
  theme(
    legend.position = "bottom",
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 7),
    legend.title = element_blank(),
    text = element_text(family = "serif")
    #panel.border = element_rect(color = "grey", fill =  NA, size = 0.5)
  )

ggsave("./images/dll_ent.pdf", device="pdf", width = 8.5, height = 4)


```



### Analysis for all languages

Analysis for all data with random by-language effects

```{r}

regression_names = c("add", "replace", "baseline")

df_all_langs = data.frame()
for (lang in langs) {
  df = read.csv(paste0("../data/merged_data/", lang, ".csv")) %>%
    mutate(lang = as.factor(lang))
  df_all_langs = rbind(df_all_langs, df)
}

#Scale things due to convergence issues w/ the random effects
#df_all_langs = df_all_langs %>%
  #group_by(model) %>%
    #mutate(surp = scale(surp), len = scale(len), freq = scale(freq), gaze_rt = scale(gaze_rt), firstfix_rt = scale(firstfix_rt), total_rt = scale(total_rt)) %>%
  #ungroup()
df_all_langs= df_all_langs[sample(1:nrow(df_all_langs)), ]
  

dll_raw_df = data.frame()
for (m in models){
for (psychometric in c("gaze_rt")) {
    
  regression_forms = c(
    paste0(psychometric, " ~ ent + prev_ent + surp + prev_surp + freq + len + prev_freq + prev_len + (ent + surp + freq + len | lang)"),
    paste0(psychometric, " ~ ent + prev_ent + freq + len + prev_freq + prev_len + (ent + freq + len | lang)"),
    paste0(psychometric, " ~ surp + prev_surp + freq + len + prev_freq + prev_len + (surp + freq + len | lang)")
  )
  
  to_fit_df = df_all_langs %>% filter(model == m) %>% drop_na()
  
  loglik_df = data.frame(names=regression_names, forms=regression_forms) %>%
    mutate(logliks = map(regression_forms, model_cross_val, df=to_fit_df, d_var=psychometric, mixed_effects=T )) %>%
    dplyr::select(-forms)
  
  loglik_df = loglik_df %>% unnest(cols = c(logliks)) %>% mutate(psychometric = psychometric, model = m)
  dll_raw_df = rbind(dll_raw_df, loglik_df)
}
}

target_df = dll_raw_df %>% filter(names == "add", model == "monot_all")
baseline_df = dll_raw_df %>% filter(names == "baseline", model == "monot_all")
dll = target_df$logliks - baseline_df$logliks
dll = dll[!is.na(dll)]
mean(dll)
test = perm.test(dll, num.sim = 1000)
test

target_df = dll_raw_df %>% filter(names == "replace", model == "monot_all")
baseline_df = dll_raw_df %>% filter(names == "baseline", model == "monot_all")
dll = target_df$logliks - baseline_df$logliks
dll = dll[!is.na(dll)]
mean(dll)
test = perm.test(dll, num.sim = 1000)
test

```


## Comparison with Random Effects by language for entropy

```{r}

regression_names = c("add", "replace", "baseline") # There are flipped from the other test where "baselines" had surprisal and "targets" didn't, confusingly.

df_all_langs = data.frame()
for (lang in langs) {

  df = read.csv(paste0("./merged_data/l1/", lang, "_clean_data.csv")) %>%
    filter(context == "long") %>%
    mutate(lang = as.factor(lang))
  df_all_langs = rbind(df_all_langs, df)
}

#Scale things due to convergence issues w/ the random effects
df_all_langs = df_all_langs %>%
  mutate(surp = scale(surp), len = scale(len), freq = scale(freq), ent = scale(ent), prev_ent = scale(prev_ent), prev2_ent = scale(prev2_ent), gaze_rt = scale(gaze_rt), firstfix_rt = scale(firstfix_rt), total_rt = scale(total_rt))
df_all_langs= df_all_langs[sample(1:nrow(df_all_langs)), ]


dll_raw_df = data.frame()
for (psychometric in c("gaze_rt")) {
    
  regression_forms = c(
    paste0(psychometric, " ~ surp + prev_surp + prev2_surp + ent + prev_ent + prev2_ent + freq*len + prev_freq*prev_len + prev2_freq*prev2_len + (1 + freq + len || lang)"),
    paste0(psychometric, " ~ ent + prev_ent + prev2_ent + freq*len + prev_freq*prev_len + prev2_freq*prev2_len + (1 + freq + len  || lang)"),
    paste0(psychometric, " ~ surp + prev_surp + prev2_surp + freq*len + prev_freq*prev_len + prev2_freq*prev2_len + (1 + freq + len || lang)")
  )
  
  loglik_df = data.frame(names=regression_names, forms=regression_forms) %>%
    mutate(logliks = map(regression_forms, model_cross_val, df=df_all_langs, d_var=psychometric, mixed_effects=T )) %>%
    dplyr::select(-forms)
  
  loglik_df = loglik_df %>% unnest(cols = c(logliks)) %>% mutate(psychometric = psychometric)
  dll_raw_df = rbind(dll_raw_df, loglik_df)
}

baseline_df = dll_raw_df %>% filter(names == "baseline")
add_df = dll_raw_df %>% filter(names == "add")
replace_df = dll_raw_df %>% filter(names == "replace")

dll_add = add_df$logliks - baseline_df$logliks
dll_add = dll[!is.na(dll)]
perm.test(dll_add, num.sim = 10000)

dll_replace = replace_df$logliks - baseline_df$logliks
dll_replace = dll[!is.na(dll)]
perm.test(dll_replace, num.sim = 10000)

```








